---
title: "Homework 1"
author: "Gaurav Kandlikar & Camila Madeiros"
date: "October 8, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 6, scipen = 999)
library(R2jags)
#setwd("/Users/camilamedeiros/Documents/PhD/Coursework/M234-Bayesian/HW_1")
setwd("~/grad/courses/UCLA/biost234/homework/hw1/")
```

## Question 1
\textsc{Explain what your measurements will be.}  
We will be exploring Gaurav's running speeds from his last ten runs.  

## Question 2  
\textsc{Deciding on some priors}  
We estimate a mean speed ($\mu_{0}$) of 8 minutes and 10 seconds (i.e. 490 seconds) per mile. The basis for this is that on good days, I (Gaurav) generally try to run 8-minute (480 seconds) miles, and I know that over the past ~2 weeks I've been running slower due to a weird knee. We estimate a prior standard deviation ($\tau$) of 30 seconds because I've done both road runs and treadmill runs, and I know that I run somewhat differently in the two conditions.  

## Question 3  
\textsc{Report the data and sample mean and variance (n-1) denominator.}  

```{r}
runs <- read.csv("running_data.csv")
head(runs)
sample_mean <- mean(runs$speed_sec_per_mi)

# Calculate sample variance
sample_var <- sum((runs$speed_sec_per_mi-sample_mean)^2)/(nrow(runs)-1) 
```

The mean of our sample is `r sample_mean`; the variance of our sample is `r sample_var`.  

## Question 4  
\textsc{Now specify the sampling standard deviation. Since we are doing a one parameter model, and since this value is usually not known, we need to do something because we are working with such a simple model.}  

We know that the speed estimates from the run tracking app are fairly accurate: the speeds from the app have closely matched my race speeds recorded independently. We don't think that the sampling standard deviation is higher than the sample $\sigma$ of `r sqrt(sample_var)`, so we will proceed in the analysis assuming the sampling standard deviation is the same as the sd of the data .

## Question 5  
\textsc{Calculate the posterior mean, variance, and SD.}  

Because we are working with a normal-normal model (mean is normally distributed with, in this case, a known sampling error), we can use the following formulae to collect posterior mean and variance:  

$$\bar{\theta} = \frac{\frac{n}{\sigma^2}}{\frac{n}{\sigma^2}+\frac{1}{\tau^2}}\bar{y} + \frac{\frac{1}{\tau^2}}{\frac{n}{\sigma^2}+\frac{1}{\tau^2}}\theta_0$$

$$\text{var}(\theta \mid \text{data}) = \left(\frac{n}{\sigma^2}+\frac{1}{\tau^2}\right)^{-1}$$

```{r}
prior_mean = 490 #seconds
prior_sd   = 30 #seconds
sampling_sd = sqrt(sample_var)
n = nrow(runs)
ybar = sample_mean

# calculate the posterior mean using the formula
posterior_mean <- (n/(sampling_sd^2))/((n/sampling_sd^2)+(1/prior_sd^2)) * ybar +
  (1/prior_sd^2)/((n/sampling_sd^2)+(1/prior_sd^2)) * prior_mean

# calculate the posterior variance using the formula
posterior_var <- ((n/(sampling_sd^2))+(1/(prior_sd^2)))^-1
posterior_sd <- sqrt(posterior_var)
```

The posterior mean is $\bar{\mu}$ = `r posterior_mean`;  
The posterior var is $V$ = `r posterior_var`;  
The posterior sd is $sd$ = `r posterior_sd`.  

\clearpage

## Question 6  
\textsc{The prior predictive density is the density that you predict for a single observation before seeing any data.}  
It sure is!  
```{r, fig.height = 5}
# Sample from a normal distribution with the parameters above
plot(density(rnorm(1000, prior_mean, sqrt(prior_sd^2+sampling_sd^2))),
     main = "prior predictive density", xlab = "speed (seconds per mile)")
```

\clearpage

## Question 7  
\textsc{Construct a table with means, sds and vars for the (i) posterior for mu, (ii) the prior for mu, (iii) the prior predictive for y, and (iv) the likelihood of mu.}  

```{r, echo = F}
posterior_row  <- c(posterior_mean, posterior_sd, posterior_var)
prior_row      <- c(prior_mean, prior_sd, prior_sd^2)
prior_pred_row <- c(prior_mean, sqrt(prior_sd^2+sampling_sd^2), 
                    prior_sd^2+sampling_sd^2) 
likelihood_row <- c(sample_mean, sqrt(sample_var/n), sample_var/n)

to_print <- rbind(posterior_row, prior_row, prior_pred_row, likelihood_row)
rownames(to_print) <- c("Posterior", "Prior", "Prior Predictive", "Likelihood")
colnames(to_print) <- c("Mean", "SD", "Variance")
knitr::kable(to_print, caption = "Posterior, Prior, Prior Predictive, and Likelihood calculated from their definitions")
```

*note: code available in appendix*

\clearpage

## Question 8  
\textsc{Plot on a single plot the (i) posterior for mu, (ii) the prior for mu, (iii) the prior predictive for y, and (iv) the likelihood of mu (suitably normalized so it looks like a density, ie a normal with mean y-bar and variance sigma2 /n) all on the same graph. Interpret the plot.}

```{r, echo = F}

posterior_vec <- rnorm(10000, posterior_mean, posterior_sd)
prior_vec <- rnorm(10000, prior_mean, prior_sd)
prior_pred_vec <- rnorm(10000, prior_mean, sqrt(prior_sd^2+sampling_sd^2))
likelihood_vec <- rnorm(10000, sample_mean, sqrt(sample_var/n))

plot(density(prior_vec), lty = 2, col = "red", main = "Probability densities from definitions",
     ylim = c(0, 0.034), lwd = 2, xlim = c(350, 600),
     xlab = "speed (seconds per mile)")
lines(density(posterior_vec), lwd = 2)
lines(density(prior_pred_vec), col = "blue", lty = 3, lwd = 2)
lines(density(likelihood_vec), col = "darkgreen", lty = 4, lwd = 2)
legend("topleft", lty = c(2, 3, 4, 1), col = c("red", "blue", "darkgreen", "black"), lwd = 2, 
       legend = c("Prior", "Prior Predictive", "Likelihood", "Posterior"), bty = "n")
```

**Interpretation of plot**:  The posterior distribution of mean running speed is in between the prior mean and the likelihood estimate; in other words, the posterior is a compromise between the likelihood and the prior, which shrank the likelihood. The prior predictive has a mean equal to the prior mean but has a much wider distribution, which comes from our uncertainty in measurements as well as the prior estimate for variation in running speed. Our posterior distribution is quite similar to the likelihood distribution, which happens because the weight given to our prior mean is lower than the weight given to the sample mean in the posterior calculation. Practically, this means that I have been running much slower than I thought over the past three weeks or so- my average speed has been about twenty seconds a mile slower than I thought. Time to fix that! 

*note: code available in appendix*


\clearpage

## Question 9  

\textsc{Write R/WinBUGS programs to sample from the posterior of mu.}

The following `JAGS` and `R` code sampels from the posterior of mu:

```{r, eval = F}
#sink("running_model.txt")
cat("model {
	for (i in 1:N) {
		x[i] ~ dnorm(mu, tau)
	}
	mu ~ dnorm(prior_mean, prior_tau) 
	sigma <- sampling_sd              
	tau <- 1/(sigma^2)             # tau equal to 1/sigma^2
}", fill = TRUE)
#sink()
```

```{r write the model}
# parameters
jags.params = c("mu", "sigma", "tau")

# data
x = runs$speed_sec_per_mi
N = length(runs$speed_sec_per_mi)
prior_tau = 1/(prior_sd^2)
jags.data = list("x", "N", "prior_mean", "sampling_sd", "prior_tau") 

# initials
jags.inits = function(){
    list("mu" = 0) # part of algorithm!
}
```

Now that we have set up the model, data, and parameters, we can run the model:
```{r run the model, eval = T, results = "hide"}
#sampling from the posterior
hw1.sim = jags(jags.data, jags.inits, jags.params, 
              model.file = "running_model.txt", 
              n.chains = 3, n.iter = 11000, n.burnin = 1000)
```

We can summarize the output of the model:

```{r}
to_print <- hw1.sim$BUGSoutput$summary[2:4,c("mean", "sd", "2.5%", "97.5%")]
knitr::kable(to_print, caption = "Summary of posterior from JAGS run")
```

## Question 10  
\textsc{Adapt your BUGS program to sample from the prior and prior predictive. Do this by not loading your data, rather, in loading the initial values, move the data y over to the init list instead. There is an example at the end of Homework 2 for a Poisson-gamma likelihood/prior. [Helpful step: set keyword DIC=F in the call to bugs, as WinBUGS can not calculate DIC for prior predictions.]}  


**Sampling from the prior**:
```{r, results = "hide"}
### Sampling from the prior:

# data
prior_tau = 1/(prior_sd)^2 # same prior tau as before
# Note that the data object 'x' is not part of the data list
jags.data.prior = list("N", "prior_mean", "sampling_sd", "prior_tau") 

# initials; note that x IS part of the inits
jags.inits_prior = function(){
    list("mu"=0, "x"=x) # part of algorithm!
}

hw1.sim2 = jags(jags.data.prior, jags.inits_prior, jags.params, 
              model.file = "running_model.txt", 
              n.chains = 3, n.iter = 11000, n.burnin = 1000, DIC=F)
```

And now, we **sample from the prior predictive**:

```{r, results = "hide"}
### Sampling from prior predictive

# data
prior_tau = 1/((prior_sd^2)+(sampling_sd^2)) # prior tau updated to be the prior predictive
# Note that the data object 'x' is not part of the data list
jags.data_prior_predictive = list("N", "prior_mean", "sampling_sd", "prior_tau") 

# initials; note that x is part of the inits
jags.inits_prior_predictive = function(){
    list("mu"= 0, "x"=x) # part of algorithm!
}

hw1.sim3 = jags(jags.data_prior_predictive, jags.inits_prior_predictive, jags.params, 
              model.file = "running_model.txt", 
              n.chains = 3, n.iter = 11000, n.burnin = 1000,
              DIC=F)
```

\clearpage

```{r, results="markup", echo = F}
to_print_prior <- hw1.sim2$BUGSoutput$summary[,c("mean", "sd", "2.5%", "97.5%")]
to_print_prior_predictive <- hw1.sim3$BUGSoutput$summary[,c("mean", "sd", "2.5%", "97.5%")]
knitr::kable(to_print_prior, caption = "Summary of prior from JAGS run")
knitr::kable(to_print_prior_predictive,  caption = "Summary of prior predictive from JAGS run")
```

## Question 11   
\textsc{Adapt your BUGS program to sample from the likelihood.}  

```{r, results = "hide"}
### Sampling from likelihood

# data
prior_tau = 1/(((prior_sd^2)/N))^2 # variance of the likelihood used here
# Note that the data object 'x' IS part of the data list:
jags.data_likelihood = list("x", "N", "prior_mean", "sampling_sd", "prior_tau") 

# initials; note that X is NO LONGER part of inits
jags.inits = function(){
    list("mu"= 0)
}

hw1.sim4 = jags(jags.data_likelihood, jags.inits, jags.params, 
              model.file = "running_model.txt", 
              n.chains = 3, n.iter = 11000, n.burnin = 1000)
    

to_print_likelihood <- hw1.sim4$BUGSoutput$summary[,c("mean", "sd", "2.5%", "97.5%")]
knitr::kable(to_print_likelihood, caption = "Summary of likelihood from JAGS run")
```

## Question 12   
\textsc{Report your WinBUGS models and R code, data, and inits. Use at least samples of size 10000.}   

\textbf{R code, model, data, and inits are reported in questions 9-11.}

\clearpage

## Question 13  
\textsc{Construct a table with means, sds and vars for the (i) posterior for mu,(ii) the prior for mu, (iii) the prior predictive for y, and (iv) the likelihood of mu from the WinBUGS output.}

```{r, echo = F}
Posterior_mu   <- hw1.sim$BUGSoutput$summary[2, c("mean", "sd")]
Prior_mu       <- hw1.sim2$BUGSoutput$summary[1,c("mean", "sd")]
Prior_predictive_y <- hw1.sim3$BUGSoutput$summary[1,c("mean", "sd")]
Likelihood_mu <- hw1.sim4$BUGSoutput$summary[2,c("mean", "sd")]

to_print <- rbind(Posterior_mu, Prior_mu, Prior_predictive_y, Likelihood_mu)

var <- as.matrix((to_print[,2])^2)

results <- cbind(to_print, var)
rownames(results) <- c("Posterior", "Prior", "Prior predictive", "Likelihood")
colnames(results) <- c("Mean", "SD", "Variance")
knitr::kable(results, caption = "Posterior, Prior, Prior Predictive, and Likelihood from JAGS simulations")
```

*note: code available in appendix*

\clearpage

## Question 14  
\textsc{Plot on a single plot the (i) posterior for mu, (ii) the prior for mu (iii) the prior predictive for y, and (iv) the likelihood of mu (suitably normalized so it looks like a density, ie a normal with mean y-bar and variance sigma2/n) all on the same graph. All from the WinBUGS output. Interpret the plot.}

```{r, echo = F}
# Extract the posterior from the BUGS run
posterior = apply(hw1.sim$BUGSoutput$sims.array, 3, unlist)
posterior <- posterior[,2]
# Extract the prior from the BUGS run
prior = apply(hw1.sim2$BUGSoutput$sims.array, 3, unlist)
prior <- prior[,1]

# Extract the prior predictive from the BUGS run
prior_predictive = apply(hw1.sim3$BUGSoutput$sims.array, 3, unlist)
prior_predictive <- prior_predictive[,1]

# Extract the likelihood from the BUGS run
likelihood = apply(hw1.sim4$BUGSoutput$sims.array, 3, unlist)
likelihood <- likelihood[,2]

plot(density(posterior), main = "Probability densities from JAGS simulations", xlim= c(350, 600), ylim= c(0, 0.034), lwd = 2,
     xlab = "speed (seconds per mile)")
lines(density(prior), lty = 2, lwd = 2, col = "red")
lines(density(prior_predictive),lty = 3, lwd = 2, col = "blue")
lines(density(likelihood),lty = 4, lwd = 2, col = "darkgreen")
legend("topleft", lty = c(1, 2, 3, 4), col = c("black", "red", "blue", "darkgreen"), lwd = 2, 
       legend = c("Posterior","Prior", "Prior Predictive", "Likelihood"), bty = "n")
```

**Interpretation of the plot**: As expected this plot is identical (more or less) to the plot generated from the same distributions in Question 8. As before, we see that the posterior distribution represents something of a compromise between the prior and likelihood distributions. Practically, this means that I have been running much slower than I thought over the past three weeks or so- my average speed has been about twenty seconds a mile slower than I thought. Time to fix that! 

*note: code available in appendix*

\clearpage

\section{Appendix}

### Question 7 code
```{r, echo = T, eval = F}
posterior_row  <- c(posterior_mean, posterior_sd, posterior_var)
prior_row      <- c(prior_mean, prior_sd, prior_sd^2)
prior_pred_row <- c(prior_mean, sqrt(prior_sd^2+sampling_sd^2), 
                    prior_sd^2+sampling_sd^2) 
likelihood_row <- c(sample_mean, sqrt(sample_var/n), sample_var/n)

to_print <- rbind(posterior_row, prior_row, prior_pred_row, likelihood_row)
rownames(to_print) <- c("Posterior", "Prior", "Prior Predictive", "Likelihood")
colnames(to_print) <- c("Mean", "SD", "Variance")
knitr::kable(to_print, caption = "Posterior, Prior, Prior Predictive, and Likelihood calculated from their definitions")
```


### Question 8 code 
```{r q8appendix, echo = T, eval = F}

posterior_vec <- rnorm(10000, posterior_mean, posterior_sd)
prior_vec <- rnorm(10000, prior_mean, prior_sd)
prior_pred_vec <- rnorm(10000, prior_mean, sqrt(prior_sd^2+sampling_sd^2))
likelihood_vec <- rnorm(10000, sample_mean, sqrt(sample_var/n))

plot(density(prior_vec), lty = 2, col = "red", main = "Probability densities",
     ylim = c(0, 0.034), lwd = 2, xlim = c(350, 600),
     xlab = "seconds per mile")
lines(density(posterior_vec), lwd = 2)
lines(density(prior_pred_vec), col = "blue", lty = 3, lwd = 2)
lines(density(likelihood_vec), col = "darkgreen", lty = 4, lwd = 2)
legend("topleft", lty = c(2, 3, 4, 1), col = c("red", "blue", "darkgreen", "black"), lwd = 2, 
       legend = c("Prior", "Prior Predictive", "Likelihood", "Posterior"), bty = "n")
```


### Question 13 code
```{r q13appendix, echo = T, eval = F}
Posterior_mu   <- hw1.sim$BUGSoutput$summary[2, c("mean", "sd")]
Prior_mu       <- hw1.sim2$BUGSoutput$summary[1,c("mean", "sd")]
Prior_predictive_y <- hw1.sim3$BUGSoutput$summary[1,c("mean", "sd")]
Likelihood_mu <- hw1.sim4$BUGSoutput$summary[2,c("mean", "sd")]

to_print <- rbind(Posterior_mu, Prior_mu, Prior_predictive_y, Likelihood_mu)

var <- as.matrix((to_print[,2])^2)
results <- cbind(to_print, var)
rownames(results) <- c("Posterior", "Prior", "Prior predictive", "Likelihood")
colnames(results) <- c("Mean", "SD", "Variance")
knitr::kable(results, caption = "Posterior, Prior, Prior Predictive, and Likelihood from JAGS simulations")
```


### Question 14 code
```{r q14appendix, echo = T, eval = F}
# Extract the posterior from the BUGS run
posterior = apply(hw1.sim$BUGSoutput$sims.array, 3, unlist)
posterior <- posterior[,2]
# Extract the prior from the BUGS run
prior = apply(hw1.sim2$BUGSoutput$sims.array, 3, unlist)
prior <- prior[,1]

# Extract the prior predictive from the BUGS run
prior_predictive = apply(hw1.sim3$BUGSoutput$sims.array, 3, unlist)
prior_predictive <- prior_predictive[,1]

# Extract the likelihood from the BUGS run
likelihood = apply(hw1.sim4$BUGSoutput$sims.array, 3, unlist)
likelihood <- likelihood[,2]

plot(density(posterior), main = "Probability densities from JAGS sumilations", xlim= c(350, 600), ylim= c(0, 0.034), lwd = 2,
     xlab = "seconds per mile")
lines(density(prior), lty = 2, lwd = 2, col = "red")
lines(density(prior_predictive),lty = 3, lwd = 2, col = "blue")
lines(density(likelihood),lty = 4, lwd = 2, col = "darkgreen")
legend("topleft", lty = c(1, 2, 3, 4), col = c("black", "red", "blue", "darkgreen"), lwd = 2, 
       legend = c("Posterior","Prior", "Prior Predictive", "Likelihood"), bty = "n")
```

