---
title: "Lab 3 submission"
author: "Gaurav Kandlikar"
date: "October 13, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

```{r load-packages, echo = F, message=FALSE, include = F}
setwd("/home/gsk/grad/courses/UCLA/biost234/homework/lab3")

#LOAD NECESSARY PACKAGES
library(R2jags)
library(lattice)
#Function to help with the burn in 
load("AddBurnin.RData")
# useful function
mysummary = function(invector) {
c(mean(invector), sd(invector), quantile(invector, .025), 
	quantile(invector,.975),
	length(invector[invector>0])/length(invector))
}
```

```{r import-data, echo = F, include = F}
TraumaData = read.table("bcj97data.txt")

#Give the columns useful names 
colnames(TraumaData) <- c("death", "n",	"intercept",	"iss",	"rts", "age", "ti", "age*ti")

#first 6 observations from proto-typical cases 
head(TraumaData, n=6) # These were made up by the authors

#First we divide up the dataset into the 6 proto-typical cases and the observed data

#For the 6 proto-typical cases define Xp the design matrix, Yp the outcomes (death=1), 
# and np the number of trials. (Why are we weighing these?)
Xp = as.matrix(TraumaData[1:6,3:8])
Yp = TraumaData[1:6,1]
np = TraumaData[1:6,2] # What is this?


#note a=Yp+1 and b=np-a+2 in the corresponding beta distributions 


# define the inverse of the Xp matrix to be used to convert 
# the prior distributions on pi to distributions on the 
# regression coefficients
invXp = solve(Xp)

# For the observed data define the design matrix, outcomes, and number of trials

Xobs = as.matrix(TraumaData[7:306,3:8])
Yobs = TraumaData[7:306,1]
nobs = TraumaData[7:306,2]

# make data frames of 100 to use later
Xobs2 = as.matrix(TraumaData[7:106,3:8])
Yobs2 = TraumaData[7:106,1]
nobs2 = TraumaData[7:106,2]
```

## To Dos from Lab 3

(1) Run jags to get 50000 samples. Get plots of the priors and posteriors, and get autocorrelation plots. 

```{r, cache = T}
todo1.data=list(invXp=invXp)
todo1.inits=rep(list(list(pie=c(0.5,0.5,0.5,0.5,0.5,0.5))),5) #5 chains 
todo1.parameters = c("betas", "pie[1:6]")

#Run the JAGS model
todo1.out = jags(todo1.data, todo1.inits, todo1.parameters, "LAB3.Priors.txt", 
	n.chains=5, n.iter=51000, n.burnin=0, n.thin=2, DIC=F)
```

```{r, echo = F}
# Treat the first 1000 iterations as a burn in	
Output1 = AddBurnin(todo1.out$BUGSoutput$sims.array,burnin=1000,n.thin=2)
```

```{r full_data, echo = F, cache = T}
todo1a.data = list(x=Xobs, y=Yobs, T = 300, invXp=invXp)
todo1a.inits = rep(list(list(pie=c(0.5,0.5,0.5,0.5,0.5,0.5))),5)
todo1a.parameters = c("betas", "pie[1:6]")

todo1a.out = jags(todo1a.data, todo1a.inits, todo1a.parameters, "Lab3.Posteriors.txt", n.chains=5, n.iter=51000, n.burnin=0, n.thin=2, DIC=F)

names(todo1a.out)
```

```{r partial_data, echo = F, cache = T}

todo1b.data = list(x=Xobs2, y=Yobs2, T = 100, invXp=invXp)
todo1b.inits = rep(list(list(pie=c(0.5,0.5,0.5,0.5,0.5,0.5))),5)
todo1b.parameters = c("betas", "pie[1:6]")

todo1b.out = jags(todo1b.data, todo1b.inits, todo1b.parameters, "Lab3.Posteriors.txt", n.chains=5, n.iter=51000, n.burnin=0, n.thin=2, DIC=F)

```

```{r, echo=F}
#Treat the first 1000 iterations as a burn in	
Output2 = AddBurnin(todo1a.out$BUGSoutput$sims.array,burnin=10000,n.thin=2)
Output3 = AddBurnin(todo1b.out$BUGSoutput$sims.array,burnin=10000,n.thin=2)

```

## Assignment   
### Q1: 1. At what lags do the autocorrelations hit zero for the 6 regression coefficients? Are the beta autocorrelations better or worse than the 6 pi’s? 

The following are autocorrelation plots for $\beta$ coefficients.

```{r, echo = F}
par(mfrow = c(2,3), oma = c(0,0,4,0))
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,1], main="beta1", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,2], main="beta2", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,3], main="beta3", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,4], main="beta4", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,5], main="beta5", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,6], main="beta6", lag.max = 300)
title(outer = T, "Autocorrelation Plots for Beta estimates")
```

The autocorrelation of the six $\beta$s hit zero at different times, and act differently after hitting zero for the first time.  

 Autocorrelation of $\beta_1$ hits 0 at Lag $\approx 70$  
 Autocorrelation of $\beta_2$ hits 0 at Lag $\approx 180$  
 Autocorrelation of $\beta_3$ hits 0 at Lag $\approx 100$  
 Autocorrelation of $\beta_4$ hits 0 at Lag $\approx 70$  
 Autocorrelation of $\beta_5$ hits 0 at Lag $\approx 70$  
 Autocorrelation of $\beta_6$ hits 0 at Lag $\approx 300$  

\clearpage

The following are autocorrelation plots for $pi$ from Full Dataset run:  
```{r, echo = F}
par(mfrow=c(2,3), oma = c(0,0,4,0))
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,7], main="Pi1", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,8], main="Pi2", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,9], main="Pi3", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,10], main="Pi4", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,11], main="Pi5", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,12], main="Pi6", lag.max = 300)
title(outer = T, "Autocorrelation Plots for Pi estimates")
```

 Autocorrelation of $\pi_1$ hits 0 at Lag $\approx 100$  
 Autocorrelation of $\pi_2$ hits 0 at Lag $\approx 150$  
 Autocorrelation of $\pi_3$ hits 0 at Lag $\approx 300$  
 Autocorrelation of $\pi_4$ hits 0 at Lag $\approx 150$  
 Autocorrelation of $\pi_5$ hits 0 at Lag $\approx 150$  
 Autocorrelation of $\pi_6$ hits 0 at Lag $\approx 150$  

On the whole, the autocorrelations for the $\beta$s reach zero at a \textbf{lower lag value} than that of the pis. There are a few pi estimates (pi1, pi5, pi6) that reach zero autocorrelation very late, though that lag varies by run. 

\clearpage 

### Q2: Turn in your properly formatted table of output for the full data set, and turn in a set of the 6 plots of the prior and posterior for the betas.

The following table summarizes the posterior distribution of Betas from the full-dataset model, and the following graphs show the posterior and prior distributions.

```{r, echo = F}
names <- paste0("betas[", 1:6, "]")
to_print <- round(Output2$Burnin.Summary[names, ], 4)
colnames(to_print) = c("Mean", "SD", "2.5%", "97.5%", "P>0")
Parameter <- c("Intercept", "ISS", "RTS", "Age", "ti", "Age * ti")
to_print <- cbind(Parameter, to_print)
knitr::kable(to_print, digits = 4, caption = "Posterior of Betas from Full Dataset Run")
par(mfrow = c(2,3), oma = c(0,0,4,0))
plot(density(Output2$Burnin.sims.matrix[,1]), xlab = "beta1", main = "")
lines(density(Output1$Burnin.sims.matrix[,1]), xlim = c(-4, 4), col = "blue", lty = 2)

plot(density(Output2$Burnin.sims.matrix[,2]), xlab = "beta2", main = " ")
lines(density(Output1$Burnin.sims.matrix[,2]), col = "blue", lty = 2, xlim = c(0, .15))

plot(density(Output2$Burnin.sims.matrix[,3]), xlab = "beta3", main = "")
lines(density(Output1$Burnin.sims.matrix[,3]), col = "blue", lty = 2, xlim = c(-1.2, .6))

plot(density(Output2$Burnin.sims.matrix[,4]), xlab = "beta4", main = "")
lines(density(Output1$Burnin.sims.matrix[,4]), col = "blue", lty = 2, xlim = c(0, .13))

plot(density(Output2$Burnin.sims.matrix[,5]), xlab = "beta5", main = "")
lines(density(Output1$Burnin.sims.matrix[,5]), col = "blue", lty = 2, xlim = c(-4, 5))

plot(density(Output2$Burnin.sims.matrix[,6]), xlab = "beta6", main = "")
lines(density(Output1$Burnin.sims.matrix[,6]), col = "blue", lty = 2, xlim = c(.1, .1))
par(xpd = T)
legend("topleft", lty = c(1,2), col = c("black", "blue"), lwd = 2, legend = c("Full data  posterior", "Prior"), inset=c(0,-.7), bty = "n")

title("Posterior and Prior distributions of Betas from Full Data run", outer = T)
```

\clearpage

### Q3: Turn in the results of the TODO step 2 properly formatted and your figures nicely annotated.

The following table summarizes estimates of $\beta$s from the partial-data model only:

```{r, echo = F}
names <- paste0("betas[", 1:6, "]")
to_print <- round(Output3$Burnin.Summary[names, ], 4)
to_print <- cbind(Parameter, to_print)

knitr::kable(to_print, digits = 4, caption = "Posterior of Betas from Partial Dataset model")
```

The following density plots show the prior distribution for $\beta$s as well as the posterior distributions estimated from the full- and partial-data runs.
```{r, echo = F, fig.align="center"}
## Make Plots
#par(mfrow = c(3,2))
par(mfrow = c(2,3), oma = c(0,0,4,0))
plot(density(Output2$Burnin.sims.matrix[,1]), xlab = "beta1", main = "")
lines(density(Output3$Burnin.sims.matrix[,1]), xlab = "beta1", col = "red", lty = 3, lwd = 2)
lines(density(Output1$Burnin.sims.matrix[,1]), xlim = c(-4, 4), col = "blue", lty = 2)

plot(density(Output2$Burnin.sims.matrix[,2]), xlab = "beta2", main = "")
lines(density(Output3$Burnin.sims.matrix[,2]), col = "red", lty = 3, lwd = 2, xlim = c(0, .15))
lines(density(Output1$Burnin.sims.matrix[,2]), col = "blue", lty = 2, xlim = c(0, .15))

plot(density(Output2$Burnin.sims.matrix[,3]), xlab = "beta3", main = "")
lines(density(Output3$Burnin.sims.matrix[,3]), col = "red", lty = 3, lwd = 2, xlim = c(-1.2, .6))
lines(density(Output1$Burnin.sims.matrix[,3]), col = "blue", lty = 2, xlim = c(-1.2, .6))

plot(density(Output2$Burnin.sims.matrix[,4]), xlab = "beta4", main = "")
lines(density(Output3$Burnin.sims.matrix[,4]), col = "red", lty = 3, lwd = 2, xlim = c(0, .13))
lines(density(Output1$Burnin.sims.matrix[,4]), col = "blue", lty = 2, xlim = c(0, .13))

plot(density(Output2$Burnin.sims.matrix[,5]), xlab = "beta5", main = "")
lines(density(Output3$Burnin.sims.matrix[,5]), col = "red", lty = 3, lwd = 2, xlab = "beta5", main = "")
lines(density(Output1$Burnin.sims.matrix[,5]), col = "blue", lty = 2, xlim = c(-4, 5))

plot(density(Output2$Burnin.sims.matrix[,6]), xlab = "beta6", main = "")
lines(density(Output3$Burnin.sims.matrix[,6]), col = "red", lty = 3, lwd = 2, xlab = "beta6", main = "")
lines(density(Output1$Burnin.sims.matrix[,6]), col = "blue", lty = 2, xlim = c(.1, .1))
par(xpd = T)
legend("topleft", lty = c(1,3,2), col = c("black", "red", "blue"), lwd = c(2,2,2), legend = c("Full data  posterior", "Partial data posterior", "Prior"), inset=c(0,-.7), bty = "n")
title("Densities of Prior and Posterior with Full/Partial Data", outer = T)
```

The following table summarizes the differences in estimates of $\beta$s depending on the dataset provided to calculate the posterior:

```{r,echo = F}
beta1_means <- c(Output1$Burnin.Summary["betas[1]","mu.vect"], Output3$Burnin.Summary["betas[1]","mu.vect"], Output2$Burnin.Summary["betas[1]","mu.vect"])
beta1_sds <- c(Output1$Burnin.Summary["betas[1]","sd.vect"], Output3$Burnin.Summary["betas[1]","sd.vect"], Output2$Burnin.Summary["betas[1]","sd.vect"])
beta2_means <- c(Output1$Burnin.Summary["betas[2]","mu.vect"], Output3$Burnin.Summary["betas[2]","mu.vect"], Output2$Burnin.Summary["betas[2]","mu.vect"])
beta2_sds <- c(Output1$Burnin.Summary["betas[2]","sd.vect"], Output3$Burnin.Summary["betas[2]","sd.vect"], Output2$Burnin.Summary["betas[2]","sd.vect"])


beta3_means <- c(Output1$Burnin.Summary["betas[3]","mu.vect"], Output3$Burnin.Summary["betas[3]","mu.vect"], Output2$Burnin.Summary["betas[3]","mu.vect"])
beta3_sds <- c(Output1$Burnin.Summary["betas[3]","sd.vect"], Output3$Burnin.Summary["betas[3]","sd.vect"], Output2$Burnin.Summary["betas[3]","sd.vect"])

beta4_means <- c(Output1$Burnin.Summary["betas[4]","mu.vect"], Output3$Burnin.Summary["betas[4]","mu.vect"], Output2$Burnin.Summary["betas[4]","mu.vect"])
beta4_sds <- c(Output1$Burnin.Summary["betas[4]","sd.vect"], Output3$Burnin.Summary["betas[4]","sd.vect"], Output2$Burnin.Summary["betas[4]","sd.vect"])

beta5_means <- c(Output1$Burnin.Summary["betas[5]","mu.vect"], Output3$Burnin.Summary["betas[5]","mu.vect"], Output2$Burnin.Summary["betas[5]","mu.vect"])
beta5_sds <- c(Output1$Burnin.Summary["betas[5]","sd.vect"], Output3$Burnin.Summary["betas[5]","sd.vect"], Output2$Burnin.Summary["betas[5]","sd.vect"])

beta6_means <- c(Output1$Burnin.Summary["betas[6]","mu.vect"], Output3$Burnin.Summary["betas[6]","mu.vect"], Output2$Burnin.Summary["betas[6]","mu.vect"])
beta6_sds <- c(Output1$Burnin.Summary["betas[6]","sd.vect"], Output3$Burnin.Summary["betas[6]","sd.vect"], Output2$Burnin.Summary["betas[6]","sd.vect"])

b1 <- cbind(beta1_means, beta1_sds)
b2 <- cbind(beta2_means, beta2_sds)
b3 <- cbind(beta3_means, beta3_sds)
b4 <- cbind(beta4_means, beta4_sds)
b5 <- cbind(beta5_means, beta5_sds)
b6 <- cbind(beta6_means, beta6_sds)

to_print <- round(rbind(b1, b2, b3, b4, b5, b6), 4)
to_print = cbind(rep(c("Prior", "Partial", "Full"), 3), to_print)
colnames(to_print) = c("model","mean", "sd")
rownames(to_print) = c("Beta1", " ", " ", "Beta2", " ", " ", "Beta3", " ", " ", "Beta4", " ", " ", "Beta5", " ", " ", "Beta6", " ", " ")
knitr::kable(to_print, caption = "Parameter estimates from Prior, Partial Data, and Full Data models")
```
  
  
### Q4: The model tracks the parameters pi1 to pi6, what is the interpretation of these parameters once the data has been incorporated?

The posterior distribution of pi1 to pi6 represents the posterior probability that the six pseudo-cases will die. The posterior of these pi estimates incorporate the expert opinion from the doctor as well as the 300 datapoints to yield a posterior. 

\clearpage 

### Q5: Extra credit: you may (but don’t need to) Turn in your answer to TODO step 3.

We add the following lines of code to our BUGS model to use the posterior predictive to make predictions about future cases:

```{r eval = F}
futurefit1 <- ilogit(inprod(people[1,],betas[]))
futurefit2 <- ilogit(inprod(people[2,],betas[]))
futurefit3 <- ilogit(inprod(people[3,],betas[]))
```

With this setup, we need to ensure that the data list that we provide JAGS include a matrix called `people` that has the same column headings as the dataset of 300, and 3 rows corresponding to the new observations. In the parameter list, we need to add `futurefit1`, `futurefit2`, and `futurefit3` so that we can retrieve these parameters after the model runs and explore their distribution.

```{r, eval = F, echo = F}
# sink("Lab3.Posteriors.Future.txt")
cat("
model{

  # Restate the priors
	pie[1]~dbeta(1.1,8.5)
	pie[2]~dbeta(3.0,11.0)
	pie[3]~dbeta(5.9,1.7)
	pie[4]~dbeta(1.3,12.9)
	pie[5]~dbeta(1.1,4.9)
	pie[6]~dbeta(1.5,5.5)

  # Convert from log odds to odds, or whatever
	for(j in 1:6){
		logitp[j]<-logit(pie[j])
	}

  	betas <- invXp %*% logitp[]

	  # NOTE: This is where the data is modeled!:

		for(i in 1:T){
		y[i] ~ dbern(p[i])
		p[i]<-ilogit(inprod(x[i,],betas[]))
		}
		
    # Using the posterior predictive to estimate survivial prob:
    futurefit1 <- ilogit(inprod(people[1,],betas[]))
    futurefit2 <- ilogit(inprod(people[2,],betas[]))
    futurefit3 <- ilogit(inprod(people[3,],betas[]))
    

}
  ",fill = TRUE)
# sink()
```

```{r lets predict bitches, echo = F, cache = T, include = F}
p1 = c(1,2,7.55,25,0,0)
p2 = c(1, 11, 7.8408, 42, 1, 42)
p3 = c(1, 16, 7.8408, 80, 1, 80)
people <- rbind(p1, p2, p3)
colnames(people) = colnames(Xobs)
prediction.data = list(x=Xobs, y=Yobs, T = 300, invXp=invXp, people = people)
prediction.inits = rep(list(list(pie=c(0.5,0.5,0.5,0.5,0.5,0.5))),5)
prediction.parameters = c("futurefit1", "futurefit2", "futurefit3" )

future.out2 = jags(prediction.data, prediction.inits, prediction.parameters, "Lab3.Posteriors.Future.txt", n.chains=5, n.iter=51000, n.burnin=0, n.thin=2, DIC=F)

FutureOut2 = AddBurnin(future.out2$BUGSoutput$sims.array,burnin=1000,n.thin=2)
```

We can plot the posterior predictive probability that each of the three patients will die and make a table of our predictions:

```{r, echo = F, fig.height=3}
par(mfrow = c(1,3), oma = c(0, 0, 4, 0))
plot(density(FutureOut2$Burnin.sims.array[,,"futurefit1"]), 
     main = "P(death) for Patient 1", xlab = "P (death)", xlim = c(0,1))
plot(density(FutureOut2$Burnin.sims.array[,,"futurefit2"]), 
     main = "P(death) for Patient 2", xlab = "P (death)", xlim = c(0,1))
plot(density(FutureOut2$Burnin.sims.array[,,"futurefit3"]), 
     main = "P(death) for Patient 3", xlab = "P (death)", xlim = c(0,1))
title("Predicted probability of death for future observations", outer = T)

to_print <- FutureOut2$Burnin.Summary
rownames(to_print) = c("Patient 1", "Patient 2", "Patient 3")
colnames(to_print) = c("Mean", "SD", "2.5%", "97.5%", "P>0")
knitr::kable(to_print, digits = 4,
             caption = "Predicted probabilities of survival for hypothetical patients")
```


\clearpage  

## Code Appendix  

### Q1

```{r, eval = F}
# Betas
par(mfrow = c(2,3), oma = c(0,0,4,0))
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,1], main="beta1", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,2], main="beta2", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,3], main="beta3", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,4], main="beta4", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,5], main="beta5", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,6], main="beta6", lag.max = 300)
title(outer = T, "Autocorrelation Plots for Beta estimates")

# Pis
par(mfrow=c(2,3), oma = c(0,0,4,0))
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,7], main="Pi1", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,8], main="Pi2", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,9], main="Pi3", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,10], main="Pi4", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,11], main="Pi5", lag.max = 300)
acf(todo1a.out$BUGSoutput$sims.array[1:5000,1,12], main="Pi6", lag.max = 300)
title(outer = T, "Autocorrelation Plots for Pi estimates")
```

### Q2 


```{r, eval = F}
names <- paste0("betas[", 1:6, "]")
to_print <- round(Output2$Burnin.Summary[names, ], 4)
colnames(to_print) = c("Mean", "SD", "2.5%", "97.5%", "P>0")
Parameter <- c("Intercept", "ISS", "RTS", "Age", "ti", "Age * ti")
to_print <- cbind(Parameter, to_print)
knitr::kable(to_print, digits = 4, caption = "Posterior of Betas from Full Dataset Run")
par(mfrow = c(2,3), oma = c(0,0,4,0))
plot(density(Output2$Burnin.sims.matrix[,1]), xlab = "beta1", main = "")
lines(density(Output1$Burnin.sims.matrix[,1]), xlim = c(-4, 4), col = "blue", lty = 2)

plot(density(Output2$Burnin.sims.matrix[,2]), xlab = "beta2", main = " ")
lines(density(Output1$Burnin.sims.matrix[,2]), col = "blue", lty = 2, xlim = c(0, .15))

plot(density(Output2$Burnin.sims.matrix[,3]), xlab = "beta3", main = "")
lines(density(Output1$Burnin.sims.matrix[,3]), col = "blue", lty = 2, xlim = c(-1.2, .6))

plot(density(Output2$Burnin.sims.matrix[,4]), xlab = "beta4", main = "")
lines(density(Output1$Burnin.sims.matrix[,4]), col = "blue", lty = 2, xlim = c(0, .13))

plot(density(Output2$Burnin.sims.matrix[,5]), xlab = "beta5", main = "")
lines(density(Output1$Burnin.sims.matrix[,5]), col = "blue", lty = 2, xlim = c(-4, 5))

plot(density(Output2$Burnin.sims.matrix[,6]), xlab = "beta6", main = "")
lines(density(Output1$Burnin.sims.matrix[,6]), col = "blue", lty = 2, xlim = c(.1, .1))
par(xpd = T)
legend("topleft", lty = c(1,2), col = c("black", "blue"), lwd = 2, 
       legend = c("Full data  posterior", "Prior"), inset=c(0,-.7), bty = "n")

title("Posterior and Prior distributions of Betas from Full Data run", outer = T)
```

### Q3 

```{r, eval = F}
names <- paste0("betas[", 1:6, "]")
to_print <- round(Output3$Burnin.Summary[names, ], 4)
to_print <- cbind(Parameter, to_print)

knitr::kable(to_print, digits = 4, caption = "Posterior of Betas from Partial Dataset model")
## Make Plots
#par(mfrow = c(3,2))
par(mfrow = c(2,3), oma = c(0,0,4,0))
plot(density(Output2$Burnin.sims.matrix[,1]), xlab = "beta1", main = "")
lines(density(Output3$Burnin.sims.matrix[,1]), xlab = "beta1", col = "red", lty = 3, lwd = 2)
lines(density(Output1$Burnin.sims.matrix[,1]), xlim = c(-4, 4), col = "blue", lty = 2)

plot(density(Output2$Burnin.sims.matrix[,2]), xlab = "beta2", main = "")
lines(density(Output3$Burnin.sims.matrix[,2]), col = "red", lty = 3, lwd = 2, xlim = c(0, .15))
lines(density(Output1$Burnin.sims.matrix[,2]), col = "blue", lty = 2, xlim = c(0, .15))

plot(density(Output2$Burnin.sims.matrix[,3]), xlab = "beta3", main = "")
lines(density(Output3$Burnin.sims.matrix[,3]), col = "red", lty = 3, lwd = 2, xlim = c(-1.2, .6))
lines(density(Output1$Burnin.sims.matrix[,3]), col = "blue", lty = 2, xlim = c(-1.2, .6))

plot(density(Output2$Burnin.sims.matrix[,4]), xlab = "beta4", main = "")
lines(density(Output3$Burnin.sims.matrix[,4]), col = "red", lty = 3, lwd = 2, xlim = c(0, .13))
lines(density(Output1$Burnin.sims.matrix[,4]), col = "blue", lty = 2, xlim = c(0, .13))

plot(density(Output2$Burnin.sims.matrix[,5]), xlab = "beta5", main = "")
lines(density(Output3$Burnin.sims.matrix[,5]), col = "red", lty = 3, lwd = 2, 
      xlab = "beta5", main = "")
lines(density(Output1$Burnin.sims.matrix[,5]), col = "blue", lty = 2, xlim = c(-4, 5))

plot(density(Output2$Burnin.sims.matrix[,6]), xlab = "beta6", main = "")
lines(density(Output3$Burnin.sims.matrix[,6]), col = "red", lty = 3, lwd = 2, 
      xlab = "beta6", main = "")
lines(density(Output1$Burnin.sims.matrix[,6]), col = "blue", lty = 2, xlim = c(.1, .1))
par(xpd = T)
legend("topleft", lty = c(1,3,2), col = c("black", "red", "blue"), lwd = c(2,2,2), 
       legend = c("Full data  posterior", "Partial data posterior", "Prior"), 
       inset=c(0,-.7), bty = "n")
title("Densities of Prior and Posterior with Full/Partial Data", outer = T)

beta1_means <- c(Output1$Burnin.Summary["betas[1]","mu.vect"],
                 Output3$Burnin.Summary["betas[1]","mu.vect"],
                 Output2$Burnin.Summary["betas[1]","mu.vect"])
beta1_sds <- c(Output1$Burnin.Summary["betas[1]","sd.vect"],
               Output3$Burnin.Summary["betas[1]","sd.vect"],
               Output2$Burnin.Summary["betas[1]","sd.vect"])
beta2_means <- c(Output1$Burnin.Summary["betas[2]","mu.vect"],
                 Output3$Burnin.Summary["betas[2]","mu.vect"],
                 Output2$Burnin.Summary["betas[2]","mu.vect"])
beta2_sds <- c(Output1$Burnin.Summary["betas[2]","sd.vect"],
               Output3$Burnin.Summary["betas[2]","sd.vect"],
               Output2$Burnin.Summary["betas[2]","sd.vect"])


beta3_means <- c(Output1$Burnin.Summary["betas[3]","mu.vect"],
                 Output3$Burnin.Summary["betas[3]","mu.vect"],
                 Output2$Burnin.Summary["betas[3]","mu.vect"])
beta3_sds <- c(Output1$Burnin.Summary["betas[3]","sd.vect"],
               Output3$Burnin.Summary["betas[3]","sd.vect"],
               Output2$Burnin.Summary["betas[3]","sd.vect"])

beta4_means <- c(Output1$Burnin.Summary["betas[4]","mu.vect"],
                 Output3$Burnin.Summary["betas[4]","mu.vect"],
                 Output2$Burnin.Summary["betas[4]","mu.vect"])
beta4_sds <- c(Output1$Burnin.Summary["betas[4]","sd.vect"],
               Output3$Burnin.Summary["betas[4]","sd.vect"],
               Output2$Burnin.Summary["betas[4]","sd.vect"])

beta5_means <- c(Output1$Burnin.Summary["betas[5]","mu.vect"],
                 Output3$Burnin.Summary["betas[5]","mu.vect"],
                 Output2$Burnin.Summary["betas[5]","mu.vect"])
beta5_sds <- c(Output1$Burnin.Summary["betas[5]","sd.vect"],
               Output3$Burnin.Summary["betas[5]","sd.vect"],
               Output2$Burnin.Summary["betas[5]","sd.vect"])

beta6_means <- c(Output1$Burnin.Summary["betas[6]","mu.vect"],
                 Output3$Burnin.Summary["betas[6]","mu.vect"],
                 Output2$Burnin.Summary["betas[6]","mu.vect"])
beta6_sds <- c(Output1$Burnin.Summary["betas[6]","sd.vect"],
               Output3$Burnin.Summary["betas[6]","sd.vect"],
               Output2$Burnin.Summary["betas[6]","sd.vect"])

b1 <- cbind(beta1_means, beta1_sds)
b2 <- cbind(beta2_means, beta2_sds)
b3 <- cbind(beta3_means, beta3_sds)
b4 <- cbind(beta4_means, beta4_sds)
b5 <- cbind(beta5_means, beta5_sds)
b6 <- cbind(beta6_means, beta6_sds)

to_print <- round(rbind(b1, b2, b3, b4, b5, b6), 4)
to_print = cbind(rep(c("Prior", "Partial", "Full"), 3), to_print)
colnames(to_print) = c("model","mean", "sd")
rownames(to_print) = c("Beta1", " ", " ", "Beta2", " ", " ", 
                       "Beta3", " ", " ", "Beta4", " ", " ", 
                       "Beta5", " ", " ", "Beta6", " ", " ")
knitr::kable(to_print, caption = "Parameter estimates from Prior, Partial Data, and Full Data models")

```

### Q5 

```{r, eval = F}
# sink("Lab3.Posteriors.Future.txt")
cat("
model{

  # Restate the priors
	pie[1]~dbeta(1.1,8.5)
	pie[2]~dbeta(3.0,11.0)
	pie[3]~dbeta(5.9,1.7)
	pie[4]~dbeta(1.3,12.9)
	pie[5]~dbeta(1.1,4.9)
	pie[6]~dbeta(1.5,5.5)

  # Convert from log odds to odds, or whatever
	for(j in 1:6){
		logitp[j]<-logit(pie[j])
	}

  	betas <- invXp %*% logitp[]

	  # NOTE: This is where the data is modeled!:

		for(i in 1:T){
		y[i] ~ dbern(p[i])
		p[i]<-ilogit(inprod(x[i,],betas[]))
		}
		
    # Using the posterior predictive to estimate survivial prob:
    futurefit1 <- ilogit(inprod(people[1,],betas[]))
    futurefit2 <- ilogit(inprod(people[2,],betas[]))
    futurefit3 <- ilogit(inprod(people[3,],betas[]))
    

}
  ",fill = TRUE)
# sink()

p1 = c(1,2,7.55,25,0,0)
p2 = c(1, 11, 7.8408, 42, 1, 42)
p3 = c(1, 16, 7.8408, 80, 1, 80)
people <- rbind(p1, p2, p3)
colnames(people) = colnames(Xobs)
prediction.data = list(x=Xobs, y=Yobs, T = 300, invXp=invXp, people = people)
prediction.inits = rep(list(list(pie=c(0.5,0.5,0.5,0.5,0.5,0.5))),5)
prediction.parameters = c("futurefit1", "futurefit2", "futurefit3" )

future.out2 = jags(prediction.data, prediction.inits, prediction.parameters, "Lab3.Posteriors.Future.txt", n.chains=5, n.iter=51000, n.burnin=0, n.thin=2, DIC=F)

FutureOut2 = AddBurnin(future.out2$BUGSoutput$sims.array,burnin=1000,n.thin=2)
par(mfrow = c(1,3), oma = c(0, 0, 4, 0))
plot(density(FutureOut2$Burnin.sims.array[,,"futurefit1"]), 
     main = "P(death) for Patient 1", xlab = "P (death)", xlim = c(0,1))
plot(density(FutureOut2$Burnin.sims.array[,,"futurefit2"]), 
     main = "P(death) for Patient 2", xlab = "P (death)", xlim = c(0,1))
plot(density(FutureOut2$Burnin.sims.array[,,"futurefit3"]), 
     main = "P(death) for Patient 3", xlab = "P (death)", xlim = c(0,1))
title("Predicted probability of death for future observations", outer = T)

to_print <- FutureOut2$Burnin.Summary
rownames(to_print) = c("Patient 1", "Patient 2", "Patient 3")
colnames(to_print) = c("Mean", "SD", "2.5%", "97.5%", "P>0")
knitr::kable(to_print, digits = 4,
             caption = "Predicted probabilities of survival for hypothetical patients")

```